Mining of Keystroke and Mouse Dynamics to Increase the Engagement of Students with Programming Assignments

### 2018-04-21 paper.tex
Reviewing corrections. Closes #12. Reviews changes. Closes #11

The proposed method in this work obtained
satisfactory results. It is usual for a classification method based on free-text
to obtain an accuracy and $\kappa$ measure below their counterparts of
fixed-text; however, in this
work the results obtained were similar to those works applied to fixed-text
dynamics. A possible explanation of this would be the addition of the mouse
dynamics features and the additional preprocessing performed on the
feature vectors.  Out of all methods tested, tree induction algorithms
obtained better results when we take in to consideration both accuracy and  $\kappa$ values.

While these are promising results further experiments are needed. Other experiments should focus on novice programmers, but that is left as future
work.

Further experiments with more data and more representative examples for all  the
clasess should give researchers the hability to do stratified samples to balence
the class distribution and achieve better results. Also more experiments should
give more insigth on the relationship between affective state  and interactive
behavior. Results show near 80\% accuracy, these results  are competitive
against other classifiers of a similar domain, but we do not know yet what is
the effectiveness when applied to affective-aware learing environments.

%This
% But you have to comment on how this addresses the problem. Which
% should be in the title that is not there yet - JJ
% You also have to draw a _hard_ conclusion. Is programming using free
% text better than what? Typing a fixed text?
% Is 80% effective?
%  - done  Mario 
Future work will focus on extracting other features from the learners\' interaction such as behavioral data, as proposed by Blikstein
\cite{blikstein2011using}; the parameters proposed by this author include the size of the program and changes in the code. Other authors also propose treating the use of certain keys as features e.g., the delete key because is manly used for correcting sentences.
% Also a multi-modal approach is needed. Since all these need
% a completely new approach, a new methodology will be developed to
% gather data and take measures.

>>>>

The proposed method in this work obtained
satisfactory results. It is usual for a classification method based on free-text
to obtain an accuracy and $\kappa$ measure below their counterparts of
fixed-text; however, in this
work the results obtained were similar to those works applied to fixed-text
dynamics. A possible explanation of this would be the addition of the mouse
dynamics features and the additional preprocessing performed on the
feature vectors.  Out of all methods tested, tree induction algorithms
obtained better results when we take into consideration both accuracy and  $\kappa$ values.

While these are promising results further experiments are
needed. Other experiments should focus on novice programmers, with
which we expect to gather more data and more representative examples for all the
clasess. These experiments should give researchers the hability to do stratified samples to balence
the class distribution and achieve better results. Also more experiments should
give more insigth on the relationship between affective state  and interactive
behavior. Results show near 80\% accuracy, these results  are competitive
against other classifiers of a similar domain, but we do not know yet what is
the effectiveness when applied to affective-aware learing environments.

%This
% But you have to comment on how this addresses the problem. Which
% should be in the title that is not there yet - JJ
% You also have to draw a _hard_ conclusion. Is programming using free
% text better than what? Typing a fixed text?
% Is 80% effective?
%  - done  Mario 
Future work will focus on extracting other features from the learners\' interaction such as behavioral data, as proposed by Blikstein
\cite{blikstein2011using}; the parameters proposed by this author
include the size of the program and changes in the code. Other authors
also propose treating the use of certain keys as features e.g., the
delete key because is manly used for correcting sentences. Also a multi-modal approach is needed. Since all these need a completely new approach, a new methodology will be developed to
gather data and take measures.
